---
title: "README"
author: "Bernard Duplan"
date: "7/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import Test Data
First, check to see if the data exists on the local machine.  If it does not, 
import and unzip it.
```{r Import}
fileLoc <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
dest <- "UCA HAR Dataset.zip"
if(!file.exists(dest)) {
    download.file(fileLoc, dest)
    unzip(dest)
}
```

### Assign Datasets to Variables
Data were taken in two phases, the training and testing phases, each contained
within a respective subfolder within the data directory.  Furthermore, the 
parent directory contains meta data denoting the activities performed and the 
features measured by the instrumentation.
Within each sub-folder, the data are contained in multiple space-delimited text 
files.  Extract each one into a variable.
```{r Assign Testing Data}
# Import Test Data
testDataFile  <- "./UCI HAR Dataset/test/X_test.txt"                            #File of "test" data values
testData  <- read.table(testDataFile)
testAcFile <- "./UCI HAR Dataset/test/y_test.txt"                               #File of "test" activity values (walking, standing, etc)
testActivity <- read.table(testAcFile) 
testSuFile <- "./UCI HAR Dataset/test/subject_test.txt"                         #File of "test" subject values (the person who was subjected to the test)
testSubject <- read.table(testSuFile)
```
```{r Assign Training Data}
# Import Train Data
trainDataFile <- "./UCI HAR Dataset/train/X_train.txt"                          #File of training data values
trainData <- read.table(trainDataFile)
trainAcFile <- "./UCI HAR Dataset/train/y_train.txt"                            #File of training activity values
trainActivity <- read.table(trainAcFile)
trainSuFile <- "./UCI HAR Dataset/train/subject_train.txt"                      #File of test subject values (the person who was subjected to the test)
trainSubject <- read.table(trainSuFile)
```
```{r Assign Meta Data}
# Import Features (Column Names)
featuresFile <- "./UCI HAR Dataset/features.txt"                                #File of channel/feature names
colNames <- read.table(featuresFile)
colNames <- colNames[,2]                                                        #First column is just a numeric index = 1:nrow(colNames)
colNames <- sub("BodyBody", "Body", colNames)                                   #Replace any variable names that repeat "Body" twice

# Import Activity Labels
activitiesFile <- "./UCI HAR Dataset/activity_labels.txt"                       #File of activity names
acNames <- read.table(activitiesFile)
```

### Set the column names of the datasets (Assignment Item 4)
The column names were imported in the features file.  See CodeBook.html for
descriptions of each variable/feature, including the respective units.
```{r}
colnames(testData) <- colNames
colnames(trainData) <- colNames
colnames(testActivity) <- "ActivityID"
colnames(trainActivity) <- "ActivityID"
colnames(testSubject) <- "SubjectID"
colnames(trainSubject) <- "SubjectID"
```

### Merge Data (Assignment Item 1)
For both the testing and training datasets, combine the Activity ID, Subject ID,
and test data values.  Then, merge the "test" and "train" rows together.
```{r Combine Columns}
# Combine ActivityID, SubjectID, and data values
testData <- cbind(testActivity, testSubject, testData)
trainData <- cbind(trainActivity, trainSubject, trainData)

# Merge the Test and Train datasets 
mergedData <- rbind(testData, trainData)
```

### Keep only the mean and standard deviation features (Assignment Item 2)
Use a regular expression to find features which include "mean()" or "std()"
```{r Select Features}
mergedData <- mergedData[, grepl("mean\\(\\)|std\\(\\)|ActivityID|SubjectID",   #Keeps only columns who's name includes either "mean()", "std()", Activityy ID, or SubjectID
    names(mergedData))]
```

### Make activity names descriptive (Assignment Item 3)
The dataset at this point contains a numeric ID for the activity.  This ID 
corresponds to a string in the acNames variable.  Index acNames to replace the 
numeric ID's with meaningfull strings.
```{r Convert Activity ID to Strings}
mergedData$ActivityID <- acNames[mergedData$ActivityID, 2]                      #Indexes the activity id's dataset with the activity id from the test/train data
```

### Find Averages (Assignment Item 5)
Average the values of each column for all combinations of ActivityID and 
SubjectID
```{r Get Averages, message=FALSE}
library(dplyr)                                                                  #Load the dplyr library
avgData <- mergedData %>%                                                       #Use string operators to group the data
    group_by(ActivityID, SubjectID) %>%                                         # and then find the averages
    summarize_all(mean)
write.table(avgData, row.names=FALSE, "avgDataTable.txt")                       #Write the data to a text file

```





