## Import "Test" Data
```{r}
testDataFile  <- "./UCI HAR Dataset/test/X_test.txt"                            #File of test data values
testData  <- read.table(testDataFile)
testAcFile <- "./UCI HAR Dataset/test/y_test.txt"                               #File of test activity values (walking, standing, etc)
testActivity <- read.table(testAcFile) 
testSuFile <- "./UCI HAR Dataset/test/subject_test.txt"                         #File of test subject values (the person who was subjected to the test)
testSubject <- read.table(testSuFile)
```

## Import "Train" Data
```{r}
trainDataFile <- "./UCI HAR Dataset/train/X_train.txt"                          #File of training data values
trainData <- read.table(trainDataFile)
trainAcFile <- "./UCI HAR Dataset/train/y_train.txt"                            #File of training activity values
trainActivity <- read.table(trainAcFile)
trainSuFile <- "./UCI HAR Dataset/train/subject_train.txt"                      #File of test subject values (the person who was subjected to the test)
trainSubject <- read.table(trainSuFile)
```

## Import Features (Column Names)
```{r}
featuresFile <- "./UCI HAR Dataset/features.txt"                                #File of channel/feature names
colNames <- read.table(featuresFile)                                            #Get the feature names for use as column/variable names
colNames <- colNames[,2]                                                        #First column is just a numeric index = 1:nrow(colNames)
colNames <- sub("BodyBody", "Body", colNames)                                   #Replace any variable names that repeat "Body" twice
```

## Import Activity Labels
```{r}
activitiesFile <- "./UCI HAR Dataset/activity_labels.txt"                       #File of activity names
acNames <- read.table(activitiesFile)
```

## Set the column names of the datasets
```{r}
colnames(testData) <- colNames
colnames(trainData) <- colNames
colnames(testActivity) <- "ActivityID"
colnames(trainActivity) <- "ActivityID"
colnames(testSubject) <- "SubjectID"
colnames(trainSubject) <- "SubjectID"
```

## Add the Activity ID and Subject ID to the datasets
```{r}
testData <- cbind(testActivity, testSubject, testData)
trainData <- cbind(trainActivity, trainSubject, trainData)
```

## Merge the Test and Train datasets (Assignment Item 1)
```{r}
mergedData <- rbind(testData, trainData)
```

## Keep only the mean and standard deviation features (Assignment Item 2)
```{r}
mergedData <- mergedData[, grepl("mean\\(\\)|std\\(\\)|ActivityID|SubjectID",   #Keeps only columns who's name includes either "mean()", "std()", Activityy ID, or SubjectID
    names(mergedData))]                   
```

## Make activity names descriptive (Assignment Item 3)
```{r}
mergedData$ActivityID <- acNames[mergedData$ActivityID, 2]                      #Indexes the activity id's dataset with the activity id from the test/train data
```

## Find the average of each variable for each activity and each subject (Assignment Item 5)
```{r echo=T, results='hide'}
library(dplyr)
```
```{r}
options(width = 90)
avgData <- tapply(mergedData$`tBodyAcc-mean()-X`, list(mergedData$ActivityID,   #Use tapply to get the mean of each combination of activity and subject
    mergedData$SubjectID), FUN=mean)                                    
avgData <- as.data.frame(t(avgData))                                            #Convert to a data frame
avgData$SubjectID <- rownames(avgData)                                          #Make the activity id a column
avgData <- select(avgData, SubjectID, as.character(unique(mergedData$ActivityID)))   #Rearrange the columns
print(avgData)
```







